---
title: "TOOL Event Log Analysis"
author: "Marco Stock"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Output Format

The interaction log provides a human-readable abstraction of modeling activities with the following columns:

- **`Timestamp (finished)`**: When the action was completed (action started at Timestamp - Duration)
- **Interaction**: Named, Renamed, Created, Assigned, Re-assigned, Unassigned, Connected, Re-connected, Disconnected, Connection aborted, Defined Cardinality, Changed Cardinality, Changed Data Type, Moved, Deleted
- **Type**: Element type (Entity, Attribute, Relationship, Link)
- **UUID**: Unique identifier of the element
- **Details**: Contextual details - for change events shows "old -> new" format (Renamed, Re-assigned, Re-connected, Changed Cardinality, Changed Data Type), movement delta (+X%, +Y%) for Moved, endpoints for Connected
- **`Duration (s)`**: Duration in seconds, calculated from raw events:
    - Created, Deleted, Disconnected: 0 (instantaneous)
    - Named/Renamed: from start of continuous typing sequence to the replay event
    - Moved: from start of continuous move sequence to the replay event
    - Assigned/Unassigned/Re-assigned: from start of continuous change-event sequence to the replay event
    - Connected/Re-connected: from start of continuous change-event sequence to the replay event
    - Defined/Changed Cardinality: from start of continuous typing sequence to the replay event
    - Changed Data Type: from start of continuous typing sequence to the replay event
- **`Pos (w,h)`**: Only for Created and Moved events, expressed as (width%, height%) from bottom-left corner

### Configuration Parameters

```{r config, class.source = "fold-show"}
# Threshold for merging consecutive replay events (in seconds)
t_thresh <- 5

# Threshold for duration calculation of raw entries (in seconds)
duration_gap_thresh <- 5

# Canvas dimensions for converting positions to percentages
canvas_width <- 1920
canvas_height <- 1080

```

### Load Event Log Data

```{r load-data, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)

# Load the event log CSV file
event_log <- read_csv("./input/event_log.csv")

# Filter to only include participants P01 - P08
event_log <- event_log %>%
  filter(ID %in% c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P08"))

# Add row numbers for reference
event_log <- event_log %>%
  mutate(row_num = row_number())
```

### Helper Functions

```{r helper-functions, results='hide', message=FALSE, warning=FALSE}
library(jsonlite)

# Helper function to safely parse JSON
safe_parse_json <- function(json_str) {
  tryCatch({
    fromJSON(json_str, simplifyVector = FALSE)
  }, error = function(e) {
    list()
  })
}

# Helper function to extract element name from JSON
extract_name <- function(details) {
  if (is.null(details)) return(NA_character_)
  details$attrs$languageAttributes$name$value %||%
    details$attrs$`.entity-name`$text %||%
    details$attrs$`.relationship-name`$text %||%
    details$attrs$`.attribute-name`$text %||%
    details$metaType$attributes$name$value %||%
    NA_character_
}

# Helper function to extract position from JSON
extract_position_raw <- function(details) {
  if (is.null(details) || is.null(details$position)) return(list(x = NA_real_, y = NA_real_))
  x <- details$position$x
  y <- details$position$y
  if (!is.null(x) && !is.null(y)) {
    return(list(x = x, y = y))
  }
  return(list(x = NA_real_, y = NA_real_))
}

# Helper function to convert position to percentage from bottom-left corner
position_to_percentage <- function(x, y, canvas_w = canvas_width, canvas_h = canvas_height) {
  if (is.na(x) || is.na(y)) return(NA_character_)
  x_pct <- round(x / canvas_w * 100, 1)
  y_pct <- round((canvas_h - y) / canvas_h * 100, 1)
  return(paste0("(", x_pct, "%, ", y_pct, "%)"))
}

# Helper function to calculate movement delta as percentage
movement_delta_percentage <- function(prev_x, prev_y, curr_x, curr_y, 
                                       canvas_w = canvas_width, canvas_h = canvas_height) {
  if (is.na(prev_x) || is.na(prev_y) || is.na(curr_x) || is.na(curr_y)) return(NA_character_)
  
  delta_x <- curr_x - prev_x
  delta_y <- curr_y - prev_y
  delta_x_pct <- round(delta_x / canvas_w * 100, 1)
  delta_y_pct <- round(-delta_y / canvas_h * 100, 1)
  x_sign <- if_else(delta_x_pct >= 0, "+", "")
  y_sign <- if_else(delta_y_pct >= 0, "+", "")
  
  return(paste0(x_sign, delta_x_pct, "%, ", y_sign, delta_y_pct, "%"))
}

# Helper function to extract parent (for Assigned/Unassigned)
extract_parent <- function(details) {
  if (is.null(details)) return(NA_character_)
  parent <- details$parent
  if (!is.null(parent)) return(parent)
  return(NA_character_)
}

# Helper function to format element type
format_element_type <- function(concept) {
  type_map <- c("entity_type" = "Entity", "attribute" = "Attribute", 
                "relationship_type" = "Relationship", "link" = "Link")
  if (concept %in% names(type_map)) type_map[concept] else concept
}

# Helper function to extract source/target for links
extract_link_endpoints <- function(details) {
  if (is.null(details)) return(list(source = NA_character_, target = NA_character_))
  source_id <- details$source$id
  target_id <- details$target$id
  return(list(
    source = if (!is.null(source_id)) source_id else NA_character_,
    target = if (!is.null(target_id)) target_id else NA_character_
  ))
}

# Helper function to extract link target position (x, y)
extract_link_target_position <- function(details) {
  if (is.null(details) || is.null(details$target)) {
    return(list(x = NA_real_, y = NA_real_))
  }
  tx <- details$target$x
  ty <- details$target$y
  return(list(
    x = if (!is.null(tx)) as.numeric(tx) else NA_real_,
    y = if (!is.null(ty)) as.numeric(ty) else NA_real_
  ))
}

# Helper function to extract cardinality (label text) from links
extract_cardinality <- function(details) {
  if (is.null(details)) return(NA_character_)
  
  labels <- details$labels
  if (is.null(labels) || length(labels) == 0) return(NA_character_)
  
  # Get the text from the first label
  label_text <- labels[[1]]$attrs$text$text
  
  if (!is.null(label_text) && label_text != "") {
    return(label_text)
  }
  
  return(NA_character_)
}

# Helper function to extract data type for attributes
extract_datatype <- function(details) {
  if (is.null(details)) return(NA_character_)
  
  datatype <- details$attrs$languageAttributes$dataType$value
  
  if (!is.null(datatype) && datatype != "") {
    return(datatype)
  }
  
  return(NA_character_)
}

# Helper function to calculate duration from a continuous sequence of events
calculate_sequence_duration <- function(events_ts, end_ts, gap_threshold) {
  if (length(events_ts) == 0) return(0)
  
  sorted_ts <- sort(events_ts, decreasing = TRUE)
  sequence_start <- sorted_ts[1]
  
  if (length(sorted_ts) > 1) {
    for (j in 2:length(sorted_ts)) {
      gap <- as.numeric(difftime(sorted_ts[j-1], sorted_ts[j], units = "secs"))
      if (gap > gap_threshold) break
      sequence_start <- sorted_ts[j]
    }
  }
  
  max(0, as.numeric(difftime(end_ts, sequence_start, units = "secs")))
}

# Helper function to create an empty interaction tibble with standard columns
create_empty_interaction_tibble <- function() {
  tibble(
    Timestamp = as.POSIXct(character()),
    Interaction = character(),
    Type = character(),
    UUID = character(),
    Details = character(),
    `Duration (s)` = numeric(),
    `Pos (w,h)` = character(),
    original_order = integer()
  )
}

# Helper function to parse JSON details and extract multiple fields
parse_details_and_extract <- function(data, extractors = list()) {
  if (nrow(data) == 0 || !"Details" %in% names(data)) return(data)
  
  result <- data %>%
    rowwise() %>%
    mutate(parsed_details = list(safe_parse_json(Details)))
  
  for (field_name in names(extractors)) {
    extractor_fn <- extractors[[field_name]]
    result <- result %>%
      mutate(!!field_name := extractor_fn(parsed_details))
  }
  
  result %>% ungroup()
}

# Type priority for determining primary delete in cascade deletions
type_priority_order <- c("Entity", "Relationship", "Attribute", "Link")
```

### Logic of the Interaction Log Generation

```{r create-interaction-log-function}
create_interaction_log <- function(participant_data, participant_id) {
  
  cat("Processing participant:", participant_id, "\n")
  
  # Add original order to preserve ordering for events with same timestamp
  participant_data <- participant_data %>%
    mutate(original_order = row_number())
  
  # Filter to replay operations only for main interaction detection
  replay_events <- participant_data %>%
    filter(Operation == "replay") %>%
    arrange(Model.Element.ID, Timestamp)
  
  if (nrow(replay_events) == 0) {
    cat("  No replay events found for", participant_id, "\n")
    return(tibble(
      `Timestamp (finished)` = as.POSIXct(character()),
      Interaction = character(),
      Type = character(),
      UUID = character(),
      Details = character(),
      `Duration (s)` = numeric(),
      `Pos (w, h)` = character()
    ))
  }
  
  # Parse JSON for each replay event
  replay_events <- replay_events %>%
    rowwise() %>%
    mutate(
      parsed_details = list(safe_parse_json(Details)),
      element_name = extract_name(parsed_details),
      pos_raw = list(extract_position_raw(parsed_details)),
      pos_x = pos_raw$x,
      pos_y = pos_raw$y,
      element_parent = extract_parent(parsed_details),
      link_info = list(extract_link_endpoints(parsed_details)),
      link_source = link_info$source,
      link_target = link_info$target,
      element_datatype = extract_datatype(parsed_details)
    ) %>%
    ungroup()

  # Generate interactions by comparing consecutive replay states
  interactions <- replay_events %>%
    group_by(Model.Element.ID) %>%
    mutate(
      prev_name = lag(element_name),
      prev_pos_x = lag(pos_x),
      prev_pos_y = lag(pos_y),
      prev_parent = lag(element_parent),
      prev_link_source = lag(link_source),
      prev_link_target = lag(link_target),
      prev_datatype = lag(element_datatype),
      event_order = row_number()
    ) %>%
    ungroup()
  
  # Helper: check if a name is a placeholder (e.g., "<Name of entity type>", "<Name>", etc.)
  is_placeholder <- function(name) {
    is.na(name) | grepl("^<.*>$", trimws(name))
  }
  
  # Determine interaction type for each replay event
  interactions <- interactions %>%
    mutate(
      prev_is_placeholder = is_placeholder(prev_name),
      curr_is_placeholder = is_placeholder(element_name),
      
      Interaction = case_when(
        # First occurrence of element = Created
        event_order == 1 ~ "Created",
        
        # Element got a parent = Assigned
        is.na(prev_parent) & !is.na(element_parent) ~ "Assigned",
        
        # Element lost parent = Unassigned
        !is.na(prev_parent) & is.na(element_parent) ~ "Unassigned",
        
        # Element changed parent = Re-assigned (parent changed from one to another)
        !is.na(prev_parent) & !is.na(element_parent) & 
          prev_parent != element_parent ~ "Re-assigned",
        
        # Link source or target changed = Re-connected
        Language.Concept == "link" & event_order > 1 &
          ((!is.na(prev_link_source) & !is.na(link_source) & prev_link_source != link_source) |
           (!is.na(prev_link_target) & !is.na(link_target) & prev_link_target != link_target)) ~ "Re-connected",
        
        # Named: Previous was placeholder/NA, current is a real name
        prev_is_placeholder & !curr_is_placeholder ~ "Named",
        
        # Renamed: Both previous and current are real names, but different
        !prev_is_placeholder & !curr_is_placeholder & 
          prev_name != element_name ~ "Renamed",
        
        # `Pos (w,h)` changed = Moved (comparing raw x,y coordinates)
        !is.na(prev_pos_x) & !is.na(pos_x) & 
          (prev_pos_x != pos_x | prev_pos_y != pos_y) ~ "Moved",
        
        TRUE ~ NA_character_
      )
    ) %>%
    select(-prev_is_placeholder, -curr_is_placeholder)
  
  # Handle special case: Links should be "Connected" on first appearance
  interactions <- interactions %>%
    mutate(
      Interaction = if_else(
        Language.Concept == "link" & event_order == 1,
        "Connected",
        Interaction
      )
    )
  
  interactions <- interactions %>%
    mutate(
      Type = sapply(Language.Concept, format_element_type),
      UUID = Model.Element.ID
    )
  
  # Create Details and `Pos (w,h)` columns
  interactions <- interactions %>%
    rowwise() %>%
    mutate(
      Details = case_when(
        Interaction == "Named" ~ element_name,
        Interaction == "Renamed" ~ paste0(prev_name, " -> ", element_name),
        Interaction == "Assigned" ~ element_parent,
        Interaction == "Re-assigned" ~ paste0(prev_parent, " -> ", element_parent),
        Interaction == "Connected" ~ paste0(link_source, " -> ", link_target),
        Interaction == "Re-connected" ~ paste0(
          "(", prev_link_source, " -> ", prev_link_target, ") to (",
          link_source, " -> ", link_target, ")"
        ),
        Interaction == "Changed Data Type" ~ paste0(prev_datatype, " -> ", element_datatype),
        Interaction == "Moved" ~ movement_delta_percentage(prev_pos_x, prev_pos_y, pos_x, pos_y),
        TRUE ~ NA_character_
      ),
      `Pos (w,h)` = case_when(
        Interaction %in% c("Created", "Moved") ~ position_to_percentage(pos_x, pos_y),
        TRUE ~ NA_character_
      )
    ) %>%
    ungroup()
  
  # Detect cardinality changes from link "change" events (not replay)
  link_changes <- participant_data %>%
    filter(Operation == "change" & Language.Concept == "link") %>%
    arrange(Model.Element.ID, Timestamp)
  
  cardinality_events <- create_empty_interaction_tibble()
  
  if (nrow(link_changes) > 0) {
    # Parse cardinality (label text) from each change event
    link_changes <- parse_details_and_extract(link_changes, list(
      cardinality = extract_cardinality
    ))
    
    # Compare consecutive states to find cardinality changes
    link_changes <- link_changes %>%
      group_by(Model.Element.ID) %>%
      mutate(
        prev_cardinality = lag(cardinality),
        event_order = row_number()
      ) %>%
      ungroup()
    
    # Identify cardinality interactions
    cardinality_events <- link_changes %>%
      mutate(
        Interaction = case_when(
          is.na(prev_cardinality) & !is.na(cardinality) ~ "Defined Cardinality",
          !is.na(prev_cardinality) & !is.na(cardinality) &
            prev_cardinality != cardinality ~ "Changed Cardinality",
          TRUE ~ NA_character_
        )
      ) %>%
      filter(!is.na(Interaction)) %>%
      mutate(
        Type = "Link",
        UUID = Model.Element.ID,
        Details = case_when(
          Interaction == "Defined Cardinality" ~ cardinality,
          Interaction == "Changed Cardinality" ~ paste0(prev_cardinality, " -> ", cardinality),
          TRUE ~ NA_character_
        ),
        `Pos (w,h)` = NA_character_
      )
    
    # Calculate duration using the helper function
    cardinality_durations <- sapply(seq_len(nrow(cardinality_events)), function(i) {
      event <- cardinality_events[i, ]
      relevant_ts <- link_changes %>%
        filter(Model.Element.ID == event$UUID, Timestamp <= event$Timestamp) %>%
        pull(Timestamp)
      calculate_sequence_duration(relevant_ts, event$Timestamp, duration_gap_thresh)
    })
    
    cardinality_events$`Duration (s)` <- cardinality_durations
    
    # Keep prev_cardinality and cardinality for merge calculations
    cardinality_events <- cardinality_events %>%
      select(Timestamp, Interaction, Type, UUID, Details, `Duration (s)`, `Pos (w,h)`,
             prev_cardinality, cardinality, original_order)
  }
  
  attribute_replays <- participant_data %>%
    filter(Operation == "replay" & Language.Concept == "attribute") %>%
    arrange(Model.Element.ID, Timestamp)
  
  datatype_events <- create_empty_interaction_tibble()
  
  if (nrow(attribute_replays) > 0) {
    # Parse data type from each replay event
    attribute_replays <- parse_details_and_extract(attribute_replays, list(
      datatype = extract_datatype
    ))
    
    # Get previous replay timestamp and datatype for each replay
    attribute_replays <- attribute_replays %>%
      group_by(Model.Element.ID) %>%
      mutate(
        prev_datatype = coalesce(lag(datatype), "String"),
        prev_replay_ts = lag(Timestamp),
        event_order = row_number()
      ) %>%
      ungroup()
    
    # Get all attribute change events with data types (parse once, reuse for detection and duration)
    attribute_changes <- participant_data %>%
      filter(Operation == "change" & Language.Concept == "attribute") %>%
      arrange(Model.Element.ID, Timestamp)
    
    attribute_changes <- parse_details_and_extract(attribute_changes, list(
      change_datatype = extract_datatype
    ))
    
    # Mark change events where data type value actually changed (typing activity)
    attribute_changes <- attribute_changes %>%
      group_by(Model.Element.ID) %>%
      mutate(
        prev_change_datatype = lag(change_datatype),
        is_datatype_typing = !is.na(change_datatype) & 
          (is.na(prev_change_datatype) | change_datatype != prev_change_datatype)
      ) %>%
      ungroup()
    
    # For each replay, check if there was data type typing activity since the previous
    # replay of the same UUID (or from log start if none exists).
    had_typing_flags <- sapply(seq_len(nrow(attribute_replays)), function(i) {
      replay <- attribute_replays[i, ]
      typing_events <- attribute_changes %>%
        filter(
          Model.Element.ID == replay$Model.Element.ID,
          is_datatype_typing == TRUE,
          Timestamp <= replay$Timestamp
        )
      if (!is.na(replay$prev_replay_ts)) {
        typing_events <- typing_events %>% filter(Timestamp > replay$prev_replay_ts)
      }
      nrow(typing_events) > 0
    })
    
    attribute_replays$had_datatype_typing <- had_typing_flags
    
    # Identify data type change interactions (when there was typing activity)
    datatype_events <- attribute_replays %>%
      filter(!is.na(datatype) & had_datatype_typing == TRUE) %>%
      mutate(
        Interaction = "Changed Data Type",
        Type = "Attribute",
        UUID = Model.Element.ID,
        Details = paste0(prev_datatype, " -> ", datatype),
        `Pos (w,h)` = NA_character_
      )
    
    # Calculate duration using the helper function
    if (nrow(datatype_events) > 0) {
      datatype_durations <- sapply(seq_len(nrow(datatype_events)), function(i) {
        event <- datatype_events[i, ]
        
        relevant_changes <- attribute_changes %>%
          filter(Model.Element.ID == event$UUID, 
                 Timestamp < event$Timestamp,
                 is_datatype_typing == TRUE)
        if (!is.na(event$prev_replay_ts)) {
          relevant_changes <- relevant_changes %>% filter(Timestamp > event$prev_replay_ts)
        }
        calculate_sequence_duration(relevant_changes$Timestamp, event$Timestamp, duration_gap_thresh)
      })
      
      datatype_events$`Duration (s)` <- datatype_durations
    }
    
    # Keep prev_datatype and element_datatype for merge calculations
    datatype_events <- datatype_events %>%
      rename(element_datatype = datatype) %>%
      select(Timestamp, Interaction, Type, UUID, Details, `Duration (s)`, `Pos (w,h)`,
             prev_datatype, element_datatype, original_order)
  }
  
  # Get raw change events (non-replay) with parsed details for duration calculation
  raw_events <- participant_data %>%
    filter(Operation == "change") %>%
    arrange(Model.Element.ID, Timestamp)
  
  # Parse raw events to extract name, position, and datatype for comparison
  if (nrow(raw_events) > 0) {
    raw_events <- raw_events %>%
      rowwise() %>%
      mutate(
        parsed_details = list(safe_parse_json(Details)),
        raw_name = extract_name(parsed_details),
        raw_pos = list(extract_position_raw(parsed_details)),
        raw_pos_x = raw_pos$x,
        raw_pos_y = raw_pos$y,
        raw_parent = extract_parent(parsed_details),
        raw_link_info = list(extract_link_endpoints(parsed_details)),
        raw_link_source = raw_link_info$source,
        raw_link_target = raw_link_info$target,
        raw_link_target_pos = list(extract_link_target_position(parsed_details)),
        raw_link_target_x = raw_link_target_pos$x,
        raw_link_target_y = raw_link_target_pos$y,
        raw_datatype = extract_datatype(parsed_details)
      ) %>%
      ungroup() %>%
      select(Model.Element.ID, Language.Concept, Timestamp, Operation, raw_name, raw_pos_x, raw_pos_y, 
             raw_parent, raw_link_source, raw_link_target, raw_link_target_x, raw_link_target_y, raw_datatype)
  }
  
  # Duration windows are bounded by the previous replay of the same UUID;
  # if none exists, the window starts at log start.
  interactions <- interactions %>%
    group_by(Model.Element.ID) %>%
    mutate(prev_replay_timestamp = lag(Timestamp)) %>%
    ungroup()
  
  # Apply duration calculation to each interaction
  if (nrow(raw_events) > 0) {
    durations <- numeric(nrow(interactions))
    
    for (i in seq_len(nrow(interactions))) {
      row <- interactions[i, ]
      interaction_type <- row$Interaction
      uuid <- row$Model.Element.ID
      replay_ts <- row$Timestamp
      prev_replay_ts <- row$prev_replay_timestamp
      
      # Created events always have duration 0
      if (is.na(interaction_type) || interaction_type == "Created") {
        durations[i] <- 0
        next
      }
      
      # Get raw events for this element since previous replay of same UUID.
      # If no previous replay exists, include all raw events up to replay timestamp.
      if (is.na(prev_replay_ts)) {
        relevant_raw <- raw_events %>%
          filter(Model.Element.ID == uuid,
                 Timestamp <= replay_ts) %>%
          arrange(Timestamp)
      } else {
        relevant_raw <- raw_events %>%
          filter(Model.Element.ID == uuid,
                 Timestamp > prev_replay_ts,
                 Timestamp <= replay_ts) %>%
          arrange(Timestamp)
      }
      
      if (nrow(relevant_raw) == 0) {
        durations[i] <- 0
        next
      }
      
      # Calculate duration based on interaction type using continuous sequence logic
      if (interaction_type %in% c("Named", "Renamed")) {
        name_ts <- relevant_raw %>% filter(!is.na(raw_name)) %>% pull(Timestamp)
        durations[i] <- calculate_sequence_duration(name_ts, replay_ts, duration_gap_thresh)
        
      } else if (interaction_type == "Moved") {
        pos_ts <- relevant_raw %>% 
          filter(!is.na(raw_pos_x) | !is.na(raw_pos_y)) %>% 
          pull(Timestamp)
        durations[i] <- calculate_sequence_duration(pos_ts, replay_ts, duration_gap_thresh)
        
      } else if (interaction_type == "Changed Data Type") {
        dt_ts <- relevant_raw %>% filter(!is.na(raw_datatype)) %>% pull(Timestamp)
        durations[i] <- calculate_sequence_duration(dt_ts, replay_ts, duration_gap_thresh)
        
      } else if (interaction_type %in% c("Assigned", "Unassigned", "Re-assigned")) {
        rel_ts <- relevant_raw %>%
          filter(!is.na(raw_pos_x) | !is.na(raw_pos_y)) %>%
          pull(Timestamp)
        durations[i] <- calculate_sequence_duration(rel_ts, replay_ts, duration_gap_thresh)
        
      } else if (interaction_type %in% c("Connected", "Re-connected")) {
        rel_ts <- relevant_raw %>%
          filter(!is.na(raw_link_target_x) | !is.na(raw_link_target_y)) %>%
          pull(Timestamp)
        durations[i] <- calculate_sequence_duration(rel_ts, replay_ts, duration_gap_thresh)
        
      } else {
        # No generic fallback to the first raw event
        durations[i] <- 0
      }
    }
    
    interactions$`Duration (s)` <- durations
  } else {
    interactions <- interactions %>%
      mutate(`Duration (s)` = 0)
  }
  interactions <- interactions %>%
    select(-prev_replay_timestamp)
  
  # Handle Deleted interactions (from remove operations)
  remove_events_raw <- participant_data %>%
    filter(Operation == "remove")
  
  if (nrow(remove_events_raw) > 0) {
    # Parse details and extract parent info
    remove_events_raw <- remove_events_raw %>%
      rowwise() %>%
      mutate(
        parsed_details = list(safe_parse_json(Details)),
        element_parent = extract_parent(parsed_details),
        Type = format_element_type(Language.Concept),
        UUID = Model.Element.ID
      ) %>%
      ungroup()
    
    # Add priority for cascade delete identification (lower = higher priority)
    remove_events_raw <- remove_events_raw %>%
      mutate(priority = match(Type, type_priority_order, nomatch = 999))
    
    # Process each timestamp group to find primary delete
    remove_events <- remove_events_raw %>%
      group_by(Timestamp) %>%
      group_modify(function(group_data, group_key) {
        if (nrow(group_data) == 1) {
          return(group_data %>% mutate(is_primary = TRUE, cascade_value = NA_character_))
        }
        
        # Multiple deletes at same timestamp - cascade delete
        deleted_uuids <- group_data$UUID
        group_data <- group_data %>%
          mutate(is_child_of_deleted = element_parent %in% deleted_uuids)
        
        # Find primary: prefer non-children with highest priority (lowest number)
        non_children <- group_data %>% filter(!is_child_of_deleted)
        best_priority <- if (nrow(non_children) > 0) min(non_children$priority) else min(group_data$priority)
        
        primary_idx <- which(
          (nrow(non_children) == 0 | !group_data$is_child_of_deleted) & 
          group_data$priority == best_priority
        )[1]
        
        # Mark primary and create cascade value
        group_data <- group_data %>%
          mutate(is_primary = row_number() == primary_idx, cascade_value = NA_character_)
        
        cascade_elements <- group_data %>% 
          filter(!is_primary) %>%
          mutate(element_str = paste0(Type, " ", UUID)) %>%
          pull(element_str)
        
        if (length(cascade_elements) > 0) {
          group_data$cascade_value[primary_idx] <- paste(cascade_elements, collapse = ", ")
        }
        group_data
      }) %>%
      ungroup() %>%
      filter(is_primary) %>%
      mutate(
        # Links use "Disconnected" instead of "Deleted"
        Interaction = if_else(Type == "Link", "Disconnected", "Deleted"),
        Details = cascade_value,
        `Pos (w,h)` = NA_character_,
        `Duration (s)` = 0  # Deletion is instantaneous
      ) %>%
      select(Timestamp, Interaction, Type, UUID, Details, `Duration (s)`, `Pos (w,h)`, original_order)
    
  } else {
    remove_events <- create_empty_interaction_tibble()
  }
  
  # Filter to valid interactions only and select columns
  interaction_log <- interactions %>%
    filter(!is.na(Interaction)) %>%
    select(Timestamp, Interaction, Type, UUID, Details, `Duration (s)`, `Pos (w,h)`, 
           pos_x, pos_y, prev_pos_x, prev_pos_y,
           prev_name, element_name, 
           prev_datatype, element_datatype,
           original_order) %>%
    bind_rows(remove_events) %>%
    bind_rows(cardinality_events) %>%
    bind_rows(datatype_events)
  
  # Remove zero-movement Moved events for entities when there's an attribute 
  assigned_events <- interaction_log %>%
    filter(Interaction == "Assigned") %>%
    select(Timestamp, parent_uuid = Details)  # Details contains the parent entity UUID
  
  if (nrow(assigned_events) > 0) {
    zero_move_artifacts <- interaction_log %>%
      filter(Interaction == "Moved" & Type == "Entity") %>%
      inner_join(assigned_events, by = c("Timestamp" = "Timestamp", "UUID" = "parent_uuid")) %>%
      filter(grepl("^[+-]0%,\\s*[+-]0%$", Details)) %>%
      select(Timestamp, UUID) %>%
      mutate(is_artifact = TRUE)
    
    if (nrow(zero_move_artifacts) > 0) {
      interaction_log <- interaction_log %>%
        left_join(zero_move_artifacts, by = c("Timestamp", "UUID")) %>%
        filter(is.na(is_artifact) | Interaction != "Moved") %>%
        select(-is_artifact)
    }
  }
  
  # Handle aborted connections: Connected with target NA followed by Disconnected
  aborted_connections <- interaction_log %>%
    filter(Interaction == "Connected" & grepl("-> NA$", Details)) %>%
    pull(UUID)
  
  if (length(aborted_connections) > 0) {
    disconnected_uuids <- interaction_log %>%
      filter(Interaction == "Disconnected") %>%
      pull(UUID)
    
    aborted_and_disconnected <- intersect(aborted_connections, disconnected_uuids)
    
    if (length(aborted_and_disconnected) > 0) {
      interaction_log <- interaction_log %>%
        mutate(
          Interaction = if_else(
            Interaction == "Connected" & UUID %in% aborted_and_disconnected & grepl("-> NA$", Details),
            "Connection aborted",
            Interaction
          ),
          Details = if_else(
            Interaction == "Connection aborted",
            gsub(" -> NA$", "", Details),
            Details
          )
        ) %>%
        filter(!(Interaction == "Disconnected" & UUID %in% aborted_and_disconnected))
    }
  }
  
  # Remove interactions for cascade-deleted elements at deletion timestamp
  
  delete_events <- interaction_log %>%
    filter(Interaction %in% c("Deleted", "Disconnected")) %>%
    select(Timestamp, UUID, Details)
  
  if (nrow(delete_events) > 0) {
    deleted_at_timestamp <- delete_events %>%
      rowwise() %>%
      mutate(
        cascade_uuids = if_else(
          !is.na(Details),
          list(str_extract_all(Details, "[0-9a-f]{8}-[0-9a-f]{4}-[0-9a-f]{4}-[0-9a-f]{4}-[0-9a-f]{12}")[[1]]),
          list(character(0))
        ),
        all_deleted_uuids = list(c(UUID, cascade_uuids))
      ) %>%
      ungroup() %>%
      select(Timestamp, all_deleted_uuids) %>%
      unnest(all_deleted_uuids) %>%
      rename(deleted_uuid = all_deleted_uuids)
    
    interaction_log <- interaction_log %>%
      anti_join(
        deleted_at_timestamp %>%
          rename(UUID = deleted_uuid),
        by = c("Timestamp", "UUID")
      ) %>%
      bind_rows(
        interaction_log %>%
          filter(Interaction %in% c("Deleted", "Disconnected"))
      ) %>%
      distinct()
  }
  
  # Sort the final log by timestamp before merging
  interaction_log <- interaction_log %>%
    arrange(Timestamp, original_order)
  
  # Define which interaction types can be merged together
  mergeable_types <- list(
    move = c("Moved"),
    naming = c("Named", "Renamed"),
    cardinality = c("Defined Cardinality", "Changed Cardinality"),
    datatype = c("Changed Data Type")
  )
  
  # Add row numbers and detect merge groups based on consecutive rows
  interaction_log <- interaction_log %>%
    mutate(row_id = row_number()) %>%
    mutate(
      end_time = Timestamp,
      start_time = Timestamp - seconds(`Duration (s)`),
      next_interaction = lead(Interaction),
      next_uuid = lead(UUID),
      next_timestamp = lead(Timestamp),
      next_duration = lead(`Duration (s)`),
      gap_to_next = as.numeric(difftime(
        next_timestamp - seconds(coalesce(next_duration, 0)),
        end_time,
        units = "secs"
      )),
      merge_category = case_when(
        Interaction %in% mergeable_types$move ~ "move",
        Interaction %in% mergeable_types$naming ~ "naming",
        Interaction %in% mergeable_types$cardinality ~ "cardinality",
        Interaction %in% mergeable_types$datatype ~ "datatype",
        TRUE ~ "none"
      ),
      next_merge_category = case_when(
        next_interaction %in% mergeable_types$move ~ "move",
        next_interaction %in% mergeable_types$naming ~ "naming",
        next_interaction %in% mergeable_types$cardinality ~ "cardinality",
        next_interaction %in% mergeable_types$datatype ~ "datatype",
        TRUE ~ "none"
      ),
      can_merge_with_next = !is.na(next_uuid) & 
        UUID == next_uuid & 
        merge_category == next_merge_category &
        merge_category != "none" &
        !is.na(gap_to_next) & gap_to_next < t_thresh
    )
  
  interaction_log <- interaction_log %>%
    mutate(
      prev_can_merge = lag(can_merge_with_next, default = FALSE),
      group_start = !prev_can_merge,
      merge_group = cumsum(group_start)
    )
  
  # Perform the merge aggregation
  interaction_log <- interaction_log %>%
    group_by(merge_group) %>%
    summarise(
      Timestamp = max(end_time),
      Interaction = first(Interaction),
      Type = first(Type),
      UUID = first(UUID),
      `Duration (s)` = as.numeric(difftime(
        max(end_time),
        min(start_time),
        units = "secs"
      )),
      `Pos (w,h)` = last(`Pos (w,h)`),
      merge_category = first(merge_category),
      n_merged = n(),
      first_prev_pos_x = first(prev_pos_x),
      first_prev_pos_y = first(prev_pos_y),
      last_pos_x = last(pos_x),
      last_pos_y = last(pos_y),
      first_prev_name = first(prev_name),
      last_element_name = last(element_name),
      first_prev_cardinality = first(prev_cardinality),
      last_cardinality = last(cardinality),
      first_prev_datatype = first(prev_datatype),
      last_element_datatype = last(element_datatype),
      original_value = first(Details),
      last_value = last(Details),
      original_order = first(original_order),
      .groups = "drop"
    ) %>%
    # Recalculate Details for merged events
    rowwise() %>%
    mutate(
      Details = case_when(
        n_merged == 1 ~ original_value,
        merge_category == "move" ~ movement_delta_percentage(
          first_prev_pos_x, first_prev_pos_y, last_pos_x, last_pos_y
        ),
        # Merged naming events
        merge_category == "naming" & Interaction == "Named" ~ last_element_name,
        merge_category == "naming" & Interaction == "Renamed" ~ paste0(first_prev_name, " -> ", last_element_name),
        # Merged cardinality events
        merge_category == "cardinality" & Interaction == "Defined Cardinality" ~ last_cardinality,
        merge_category == "cardinality" & Interaction == "Changed Cardinality" ~ 
          paste0(first_prev_cardinality, " -> ", last_cardinality),
        # Merged datatype events
        merge_category == "datatype" ~ paste0(first_prev_datatype, " -> ", last_element_datatype),
        # Fallback
        TRUE ~ original_value
      )
    ) %>%
    ungroup() %>%
    # Select final columns
    select(`Timestamp (finished)` = Timestamp, Interaction, Type, UUID, Details, `Duration (s)`, `Pos (w,h)`, original_order) %>%
    arrange(`Timestamp (finished)`, original_order) %>%
    select(-original_order)  # Remove internal column from final output
  
  cat("  Generated", nrow(interaction_log), "interactions\n")
  
  return(interaction_log)
}
```

### Generate Logs Per Participant

```{r generate-logs-per-participant}
participants <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P08")

output_dir <- "output"
if (!dir.exists(output_dir)) dir.create(output_dir)

interaction_summaries <- list()

for (pid in participants) {
  # Filter data for this participant
  participant_data <- event_log %>%
    filter(ID == pid)
  
  if (nrow(participant_data) == 0) {
    cat("No data found for participant:", pid, "\n\n")
    next
  }
  
  # Generate interaction log
  interaction_log <- create_interaction_log(participant_data, pid)
  
  # Save to CSV
  output_file <- file.path(output_dir, paste0("interaction_log_", pid, ".csv"))
  write_csv(interaction_log, output_file)
  cat("  Saved to:", output_file, "\n\n")
  
  # Store summary for later (count by Interaction and Type)
  interaction_summaries[[pid]] <- interaction_log %>%
    count(Interaction, Type, name = "Count") %>%
    mutate(Participant = pid)
}
```

```{r summary-all-participants, class.source = "fold-show"}
# Combine all summaries
all_summaries <- bind_rows(interaction_summaries)

# Create combined Interaction_Type column for display
summary_wide <- all_summaries %>%
  mutate(Interaction_Type = paste0(Interaction, " (", Type, ")")) %>%
  select(-Interaction, -Type) %>%
  pivot_wider(
    names_from = Participant,
    values_from = Count,
    values_fill = 0
  ) %>%
  arrange(Interaction_Type)

print(summary_wide, n = 50)
```
